\pkg{cubble}: An R Package for Structuring Spatio-temporal Data

### 2021-11-25 main.Rmd
re-structure introduction and section 2 WIP

@@

**My proposal**  
Recent software development in R has proposed several relational data structure: `tidygraph`[@tidygraph] for graph manipulation, `dm` [@dm] for relational data model, while spatio-temporal data could benefit from having its own relational data structure that identifies the essential spati0-temporal element: site identifier, time identifier, geographic coordinates. In this paper, we propose a tidy data structure for vector spaio-temporal data.  \newline

**Section division**  
The rest of the paper will be divided as follows: Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data. Section 3 presents a new data structure for spatio-temporal data: cubble. Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4. Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data. \newline

@@

**Existing packages**  
Many data structures have been proposed for spatial and temporal data, but not many for spatio-temporal data. One of the reason could be the inherent different levels of information make it inefficient to store in the same table. `spacetime`[@spacetime] proposed four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT) based on underlying spatial structure `sp` [@sp] and temporal structure `xts`[@xts]. `spatstat` [@spatstat] implements a `ppp` class for point pattern data. More recent package `stars` [@stars] uses a spatiotemporal array to store the data and the array structure has its influence from `cubelyr` [@cubelyr], a dplyr data cube backend. \newline

>>>>

**Motivation**

In spatio-temporal data analysis, there's a need for a new spato-temporal data structure. The requirement for such a tool is necessary given the ubiquity of spatio-temporal data in the wild. Example of this includes house price of a city or county, climate station data from the government. The ireland wind data from

**Existing packages**

Many data structures have been proposed for spatial and temporal data, but not many for spatio-temporal data. One of the reason could be the inherent different levels of information make it inefficient to store in the same table. `spacetime`[@spacetime] proposed four space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI), and trajectory (STT) based on underlying spatial structure `sp` [@sp] and temporal structure `xts`[@xts]. `spatstat` [@spatstat] implements a `ppp` class for point pattern data. More recent package `stars` [@stars] uses a spatiotemporal array to store the data and the array structure has its influence from `cubelyr` [@cubelyr], a dplyr data cube backend. \newline

**Our new data structure for spatio-temporal data**

In this paper, a new spatio-temporal data structure, `cubble`, is introduced. `cubble` implements a relational data structure that splits the data into its spatial and temporal component. With this structure, users can separately manipulate the spaital or temporal component of the data while keep the two dimension linked. The software is available from the Comprehensive R Archive Network (CRAN) at [CRAN link]. \newline

**Section division**

The rest of the paper will be divided as follows: Section 2 reviews the existing data structure for spatio, temporal, and spatio-temporal data. Section 3 presents a new data structure for spatio-temporal data: cubble. Then the paper introduces the workflow of data manipulation and visualisation with the cubble structure in Section 4. Section 5 gives some examples on how common spatial and temporal manipulations are performed with cubble and how static and interactive visualisation help to understand climate and [...] data. \newline

@@

**My proposal**  
Recent software development in R has proposed several relational data structure: `tidygraph`[@tidygraph] for graph manipulation, `dm` [@dm] for relational data model, while spatio-temporal data could benefit from having its own relational data structure that identifies the essential spati0-temporal element: site identifier, time identifier, geographic coordinates. \newline

@@

### 2021-11-25 main.tex
re-structure introduction and section 2 WIP

\textbf{My proposal}\\
Recent software development in R has proposed several relational data
structure: \texttt{tidygraph}\citep{tidygraph} for graph manipulation,
\texttt{dm} \citep{dm} for relational data model, while spatio-temporal
data could benefit from having its own relational data structure that
identifies the essential spati0-temporal element: site identifier, time
identifier, geographic coordinates. In this paper, we propose a tidy
data structure for vector spaio-temporal data. \newline

\textbf{Section division}\\
The rest of the paper will be divided as follows: Section 2 reviews the
existing data structure for spatio, temporal, and spatio-temporal data.
Section 3 presents a new data structure for spatio-temporal data:
cubble. Then the paper introduces the workflow of data manipulation and
visualisation with the cubble structure in Section 4. Section 5 gives
some examples on how common spatial and temporal manipulations are
performed with cubble and how static and interactive visualisation help
to understand climate and {[}\ldots{]} data. \newline

\textbf{Existing packages}\\
Many data structures have been proposed for spatial and temporal data,
but not many for spatio-temporal data. One of the reason could be the
inherent different levels of information make it inefficient to store in
the same table. \texttt{spacetime}\citep{spacetime} proposed four
space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI),
and trajectory (STT) based on underlying spatial structure \texttt{sp}
\citep{sp} and temporal structure \texttt{xts}\citep{xts}.
\texttt{spatstat} \citep{spatstat} implements a \texttt{ppp} class for
point pattern data. More recent package \texttt{stars} \citep{stars}
uses a spatiotemporal array to store the data and the array structure
has its influence from \texttt{cubelyr} \citep{cubelyr}, a dplyr data
cube backend. \newline

@@

@@

\hypertarget{workflow-data-pipeline}{%
\section{Workflow/ Data pipeline}\label{workflow-data-pipeline}}

@@

\hypertarget{compatibility-with-existing-software}{%
\subsection{Compatibility with existing
software}\label{compatibility-with-existing-software}}

>>>>

\textbf{Motivation}

In spatio-temporal data analysis, there's a need for a new
spato-temporal data structure. The requirement for such a tool is
necessary given the ubiquity of spatio-temporal data in the wild.
Example of this includes house price of a city or county, climate
station data from the government,

\textbf{Existing packages}

Many data structures have been proposed for spatial and temporal data,
but not many for spatio-temporal data. One of the reason could be the
inherent different levels of information make it inefficient to store in
the same table. \texttt{spacetime}\citep{spacetime} proposed four
space-time layouts: Full grid (STF), sparse grid(STS), irregular (STI),
and trajectory (STT) based on underlying spatial structure \texttt{sp}
\citep{sp} and temporal structure \texttt{xts}\citep{xts}.
\texttt{spatstat} \citep{spatstat} implements a \texttt{ppp} class for
point pattern data. More recent package \texttt{stars} \citep{stars}
uses a spatiotemporal array to store the data and the array structure
has its influence from \texttt{cubelyr} \citep{cubelyr}, a dplyr data
cube backend. \newline

\textbf{Our new data structure for spatio-temporal data}

In this paper, a new spatio-temporal data structure, \texttt{cubble}, is
introduced. \texttt{cubble} implements a relational data structure that
splits the data into its spatial and temporal component. With this
structure, users can separately manipulate the spaital or temporal
component of the data while keep the two dimension linked. The software
is available from the Comprehensive R Archive Network (CRAN) at {[}CRAN
link{]}. \newline

\textbf{Section division}

The rest of the paper will be divided as follows: Section 2 reviews the
existing data structure for spatio, temporal, and spatio-temporal data.
Section 3 presents a new data structure for spatio-temporal data:
cubble. Then the paper introduces the workflow of data manipulation and
visualisation with the cubble structure in Section 4. Section 5 gives
some examples on how common spatial and temporal manipulations are
performed with cubble and how static and interactive visualisation help
to understand climate and {[}\ldots{]} data. \newline

@@

\textbf{My proposal}\\
Recent software development in R has proposed several relational data
structure: \texttt{tidygraph}\citep{tidygraph} for graph manipulation,
\texttt{dm} \citep{dm} for relational data model, while spatio-temporal
data could benefit from having its own relational data structure that
identifies the essential spati0-temporal element: site identifier, time
identifier, geographic coordinates. \newline

@@

\hypertarget{cubble-data-structure}{%
\section{Cubble data structure}\label{cubble-data-structure}}

@@

\hypertarget{integrating-into-a-tidy-workflow}{%
\subsection{Integrating into a tidy
workflow}\label{integrating-into-a-tidy-workflow}}
